Manuscript figures

### 2025-07-10 Manuscript_figures.Rmd
Updated for revision

c("PLXDC2","CBLB","TCF4",
  "GPAM","ACSL1","LPL","DGAT2",
  "PRSS23","THSD7A","AL139317.5") %>% 
  sccore::dotPlot(., 
                  cm.merged, 
                  anno.adipocytes, 
                  gene.order = ., 
                  cols = c("white","red3"))
```

@@

Load data. We need the `sds_obj` from Figure 6A. Also, for ease of use we provide slingshot pseudotime W/O visit 2 as `pseudotime_novis2.qs`.

@@

```{r, fig.width = 6, fig.height = 4}
tmp <- con.integration$clusters$loft_transfer$groups %>%
  renameAnnotation("Endothelial cells", "ECs") %>%
  renameAnnotation("FAPs", "ASPCs") %>% 
  renameAnnotation("Lymphoid immune cells", "Lymphoid ICs") %>%
  renameAnnotation(c("Tissue-resident macrophages", "Myeloid immune cells"), "Myeloid ICs") %>% 
  {data.frame(cid = names(.), anno = unname(.))} %>% 
  filter(!grepl("!!", cid)) %>% 
  mutate(sample = con.integration$clusters$hinte_condition$groups[match(cid, names(con.integration$clusters$hinte_condition$groups))])

@@

df.plot <- tmp.prop %>% 
  {Map(\(nn, x) data.frame(sample = nn, anno = names(x), value = unname(x)), nn = names(.), x = .)} %>% 
  bind_rows() %>% 
  mutate(anno = factor(anno) %>% 
           factor(levels = levels(.)[c(1, 2, 5, 8, 3, 7, 6, 4)]) %>% 
           renameAnnotation("Lymphatic endothelial cells", "LECs"))

@@

@@

@@

@@

```{r s5}
con <- qread("con_adipocytes.qs", nthreads = 10)

@@

```{r s5f}
con$findCommunities(resolution = 1, test.stability = T, name = "l05stab")

plot_grid(plotlist = list(
  con$plotClusterStability("l05stab", "ari") +
    theme_bw() +
    theme(line = element_blank(),
          axis.text.x = element_blank()) + 
    geom_boxplot(aes(col = "#E41A1C"), notch = T) +
    labs(x = " \n "),
  con$plotClusterStability("l05stab", "hjc") + 
    theme_bw() +
    theme(line = element_blank()) + 
    scale_color_manual(values = RColorBrewer::brewer.pal(8, "Reds")[-1])
), ncol = 2, rel_widths = c(1.3,3))
```

@@

```{r s5g}
cm.merged <- con$getJointCountMatrix()

anno.adipocytes <- qread("anno_adipocytes_archetypes.qs")

c("PRSS23","PDE5A","ADIPOQ","DGAT2", "GPAM", "LPL", "DCN", "VIM", "RORA", "CLSTN2", "CRIM1", "ITGA1") %>% 
  sccore::dotPlot(., 
                  cm.merged, 
                  anno.adipocytes, 
                  gene.order = ., 
                  cols = c("white","red3"))
```

@@

First, we integrate adipocytes from both studies. This is not run here

@@

Load data. For ease, we also provide pseudotime W/O visit 2 nuclei. We need the `sds_obj` from Figure 6A. Also, for ease of use we provide slingshot pseudotime W/O visit 2 as `pseudotime_novis2.qs`.

@@

We need the `sds_obj` object from Figure 6A.

>>>>

# c("PLXDC2","CBLB","TCF4",
#   "GPAM","ACSL1","LPL","DGAT2",
#   "PRSS23","THSD7A","AL139317.5") %>% 
c("PRSS23","PDE5A","ADIPOQ", "DGAT2", "GPAM", "LPL", "RORA", "CLSTN2", "CRIM1", "ITGA1") %>% 
  sccore::dotPlot(., 
                  cm.merged, 
                  anno.adipocytes, 
                  gene.order = ., 
                  cols = c("white","red3"))
```

@@

Load data. We need the `sds_obj` from Figure 6a. Also, for ease of use we provide slingshot pseudotime W/O visit 2 as `pseudotime_novis2.qs`.

@@

```{r, fig.width = 6, fig.height = 4}
tmp <- loft.transfer$labels %>%
  renameAnnotation("Lymphoid immune cells", "Lymphoid ICs") %>%
  renameAnnotation(c("Myeloid immune cells"), "Myeloid ICs") %>% 
  {data.frame(cid = names(.), anno = unname(.))} %>% 
  filter(!grepl("!!", cid)) %>% 
  mutate(sample = con.integration$clusters$hinte_condition$groups[match(cid, names(con.integration$clusters$hinte_condition$groups))])

@@

df.plot <- tmp.prop %>% 
  {Map(\(nn, x) data.frame(sample = nn, anno = names(x), value = unname(x)), nn = names(.), x = .)} %>% 
  bind_rows() %>% 
  mutate(anno = factor(anno) %>% 
           factor(levels = levels(.)[c(1, 2, 5, 8, 3, 7, 6, 4, 9)]) %>% 
           renameAnnotation("Lymphatic ECs", "LECs"))

@@

We integrated our data with the data from [Hinte *et al.*](https://www.nature.com/articles/s41586-024-08165-7) that were graciously provided by the authors. As these data are not public, here we will only show how we plotted the data.

@@

```{r, fig.width=4, fig.height=4}
# Use Cacoa to calculate proportions
cao <- Cacoa$new(data.object = con.major, 
                 cell.groups = anno.major, 
                 ref.level = "vis2", 
                 target.level = "vis3", 
                 sample.groups = con.major$samples %>% 
                   names() %>% 
                   strsplit("_") %>% 
                   sapply('[[', 3) %>% 
                   gsub(1, 2, .) %>% 
                   `names<-`(con.major$samples %>% 
                               names()),
                 sample.groups.palette = ggsci::pal_jama()(7))

@@

p <- data.frame(Proportions = proportions,
           Predictions = unname(lasso.predict.all)) %>% 
  ggplot(aes(Proportions, Predictions)) + 
  geom_point() +
  theme_bw() +
  theme(axis.text = element_text(colour = "black")) +
  labs(title = paste0("r2 = ", formatC(r2.all, digits = 3)), y = "Predicted proportions (%)", x = "Proportions (%)") +
  geom_abline(intercept = 0, slope = 1, colour = "firebrick")

@@

```{r s5}
con <- qread("con_adipocytes.qs", nthreads = 10)
anno.adipocytes <- qread("anno_adipocytes_archetypes.qs")
spc <- con$getDatasetPerCell()
```

@@

```{r s5f}
con$findCommunities(resolution = 1, test.stability = T, name = "l1stab")

plot_grid(plotlist = list(
  con$plotClusterStability("l05stab", "ari") +
    theme_bw() +
    theme(line = element_blank(),
          axis.text.x = element_blank()) + 
    geom_boxplot(aes(col = "#E41A1C"), notch = T) +
    labs(x = " \n "),
  con$plotClusterStability("l1stab", "hjc") + 
    theme_bw() +
    theme(line = element_blank()) + 
    scale_color_manual(values = RColorBrewer::brewer.pal(8, "Reds")[-1])
), ncol = 2, rel_widths = c(1.3,3))
```

@@

```{r s5g, fig.width=6, fig.height=4}
cm.merged <- con$getJointCountMatrix()

c("PRSS23","PDE5A","ADIPOQ", "DGAT2", "GPAM", "LPL", "RORA", "CLSTN2", "CRIM1", "ITGA1") %>% 
  sccore::dotPlot(., 
                  cm.merged, 
                  getConosCluster(con, "leiden0.5"), 
                  gene.order = ., 
                  cols = c("white","red3"))
```

@@

First, we integrate adipocytes from both studies. This is not run here, and the data are not shared.

@@

Load data. For ease, we also provide pseudotime W/O visit 2 nuclei. We need the `sds_obj` from Figure 6a. Also, for ease of use we provide slingshot pseudotime W/O visit 2 as `pseudotime_novis2.qs`.

@@

We need the `sds_obj` object from Figure 6a.
